% -*- coding: utf-8; -*-

\chapter{Introduction}

\section{Motivation}

In nature, there are 20 different natural amino acids, each of them exhibiting different pysicochemical properties (e.g. size, charge, hydrophobicity)\cite{1}\cite{2}. Depending on their polarity, amino acids  vary in their hydrophilic or hydrophobic character, which is crucial for the formation of more complex structures. A peptide is a molecule composed of two or more amino acid residues linked through a covalent chemical bond called the peptide bond. This peptide bond is formed when the carboxyl group of a residue reacts with the amino group of another residue,releasing one water molecule. Large peptides are generally called polypeptides or proteins\cite{3}\cite{4}. Proteins perform a variety of functions in living organisms such as catalysts of specific reactions, structural components of cell membranes, antibodies, carry oxygen in the blood and are part of chromosome material. On the other hand, proteins control the regulation and reproduction of living things and are thus essential for life.%that are vital for life.Furthermore, proteins are important components of cell %membranes, function as antibodies,carry oxygen in the blood and are part of chromosome %material. Proteins control the regulation and reproduction of living things and are %thus essential for life.

The structure of proteins can be divided into four levels\cite{1}\cite{2}  (a) primary structure, (b) secondary structure, (c) tertiary structure and (d) Quaternary structure. The primary structure corresponds to the sequence (linear order) of the amino acids forming a given protein\cite{1}\cite{2}\cite{4}\cite{5}. The beginning of the primary structure corresponds to its N-terminal and the end of its primary structure is the C-terminal region. The secondary structure is defined by characteristic structural arrangements that are the consequence of hydrogen-bond patterns involving the backbone atoms (the amide chemical function that forms the main chain of the peptide)\cite{2}. The most common regular secondary structures are $\alpha$-helices and $\beta$-sheets\cite{6}, which are usually highly stable and constitute key elements in the three-dimensional (3D) structure of proteins. The tertiary structure of a protein is represented by the distribution of secondary structures in a 3D space. 
\newpage
The three-dimensional shape assumed by a protein is also called its native, functional or folded structure. The native structure is the lowest free-energy conformation under physiological condition, which can be viewed as a consequence of the compromise between enthalpy (Intra- and intermolecular interactions) and entropic (related to the number of accessible states) contributions\cite{1}\cite{2}\cite{4}\cite{7}. The knowledge of the tertiary structure of a protein is usually of great importance in biology as it permits: the analysis and prediction of protein function in cells; the identification of active sites and binding sites of a receptor; or the identification of a site of recombination to the action of another protein\cite{2}. Finally, the quaternary structure is defined for proteins having multiple domains of tertiary structure and is defined by the spatial arrangement of these domains.

One of the main challenges in structural bioinformatics concerns the determination of protein tertiary structures. Determining the tertiary structure of proteins are experimentally and computationally expensive and time consuming\cite{8}. The difficulty in determining 3D structures of proteins has generated a large imbalance between the number of known amino acid sequences and the number of 3D structures of proteins. In spite of a large number of known protein sequences, only a small fraction possess associated 3D structures. This emphasizes the need of computational methods for the prediction of protein tertiary structures.


A number of computational methods have been proposed as a solution to the problem of protein structure prediction (PSP)\cite{9}\cite{10}\cite{11}\cite{12}. These methods can be divided into four classes \cite{13}: (a) fold recognition and threading methods\cite{14}\cite{15}\cite{16}\cite{17}; (b) comparative modeling methods and sequence alignment strategies \cite{18}\cite{19} (c) first principle methods with database information\cite{20}\cite{21}; (d) first principle methods without database information\cite{10}. The first three groups of methods are capable of predicting the tertiary structure of proteins quickly and efficiently when template structures or folding libraries are known\cite{22}. However, the last group (first principle methods without database information), which is referred to as ab initio, is based on first principles and do not rely on database information and produces new folds through computer simulation of physicochemical properties of the process of protein folding in nature. All these techniques generate a large number of candidate models known as decoys \cite{23}. Determine the models quality assessment is a challenging problem in structural biology and is an important area in bioinformatics\cite{24}.Several model quality assessment methods(MQA) have been developed since CASP7 (7th Community Wide Experiment on the Critical Assessment of Techniques for Protein Structure Prediction)\cite{24}\cite{25}.

Over the past years, several techniques have been developed, these techniques can be classified into two big groups: When the  native protein structure is known and when is not. The first group of techniques evaluates the quality of a decoy through different measures of similarity such as Root-Mean-Square Deviation (RMSD)\cite{26}, Global Distance Test Total Score (GDT\_TS) \cite{27}, Template-Modeling Score (TM-Score)\cite{28} and Maximal Substructure (MaxSub)\cite{29}. The second group of methods uses scoring functions that allows one to discriminate between high and low-quality models. These scoring functions can be classified as follows: (a)physics based potential functions; (b)statistical potential functions; (c)consensus-based functions; and (d)machine learning algorithms.Physics based functions calculate the potential energy associated with a given protein model and may incorporate also the interactions with the surrounding solvent\cite{30}\cite{31}. Such methods usually consume a significant amount of computational time and are not very sensitive to small atomic changes. Statistical potential functions evaluate the quality of the  models based on statistical analysis of structural attributes extracted from a known protein structure database\cite{32}\cite{33}. However, these methods only consider average properties of known protein structures and therefore have limited power to discriminate and rank structural models. Consensus-based functions have a good performance when many models in the database are similar to the native structure. However, if the database is composed of only low-quality models, these methods tend to show less performance than knowledge-based approaches\cite{34}\cite{35}. Machine learning algorithms, such as support vector machine (SVM), neural networks (NN) and random forest (RF), evaluate the quality of the models in relation to certain features or attributes\cite{36}\cite{37}. These features are extracted from the sequence and structure of decoy models and are used as inputs to evaluate their quality. The great advantage of these methods is that they can consider a large number of features. These features may be considered simultaneously, and the hidden relationships between them can be taken into account, which is very difficult to achieve with the other methods described.
% % %\newpage

Features can be extracted from the secondary structure (for example, average of amino acids forming alpha helix structures). Features based on Euclidean distance are also widely used (for example, distance between the amino acids and solvent distance between aliphatic amino acids). Features can also be calculated based on the physicochemical properties of protein models, such as the number of aliphatic hydrophobic amino acids and the solvent accessibility on different levels\cite{38}\cite{39}\cite{40}. Furthermore, it is also possible to represent the protein as a network structure\cite{41}. Using such a representation it is possible to calculate features such as: (a)number of non covalent interactions (defined by the number of sides on the network); (b)set of nodes connected with maximum number of amino acids; (c)average cluster coefficient of the network; and (d)average cluster coefficient of the largest cluster. 

Given this wide range of possible features, many studies select an initial set to be used as inputs for the machine learning model. This selection is performed in an arbitrary fashion, which implies in biased results and, therefore, limits the prediction power of the method.

Therefore, it is highly desirable to obtain an unbiased set of relevant features in an automatic way. This could be achieved by a computational method capable of, in a first step, perform automatic feature selection identifying the most important ones from a large pool of available features and, in a second step, evaluate the relative importance of each feature within this reduced subset.

\section{Objectives}
The main objective of this study is to calculate the importance of different types of features in protein quality assessment. The specific objectives are:

\begin{enumerate}[label=(\roman{*})]
  \item To propose a model that is able to evaluate the relative importance for an optimal subset of features;
  \item To use this model to predict quality of decoy models towards their native structures. 
\end{enumerate}

\section{Contributions}
The main contribution of this work is the proposal of a strategy that uses a hybrid model, based on a genetic algorithm and a neural network, for the calculation of the characteristics (features). This model allows specifically:
\begin{enumerate}[label=(\roman{*})]
	\item Calculate three different scores for each of the features used in this proposal;
	\item Calculate the similarity of decoys models relative to its native structure. 
\end{enumerate}

\section{Work Description}
The development of this study was done in the following steps:
\begin{enumerate}[label=(\roman{*})]
	\item Study of the main technical similarity in the literature such as RMSD, TM-score, GDT\_TS, MaxSub.
	\item Study of machine learning techniques to solve this problem (Neural Networks, SVMR, Random Forest etc.). 
	\item Search of features commonly used in machine learning models described above
	\item Pre-processing decoys database, which is necessary for the feature extraction (calculation). 
	\item Hybrid model definition (GA + NN), for the calculation of features importance.
	\item Implementation (Matlab and Python) and testing of the proposed model in programming languages, to adjust the proposed model.
\end{enumerate}

\section{Work Organization}

The proposal is divided into four additional chapters. Chapter 2 presents a summary of the theoretical foundations necessary for understanding this work, which will consider issues such as amino acids, proteins and its different levels of organization, methods for predicting the 3D protein structure, protein structure similarity metrics, and machine learning methods to predict the quality of decoys. Finally, the chapter shows the different features used to predict the protein quality assessment. In Chapter 3, the model for the protein quality assessment calculation will be presented in details, emphasizing the calculation mechanisms of the importance of the features. The chapter 4 describes and analyzes the preliminary results obtained until now. Finally, Chapter 5 presents the discussions and conclusions on the proposed model and the results obtained. 